---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r echo=FALSE}
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data
Analysis of data obtained from an activity monitor for an individual that measured the steps taken over a 5 minute period over a number of days. 

The raw data was downloaded and read it into a data frame, ready for the analysis.

```{r}
url<-"https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url,"activity.zip")
unzip("activity.zip")
activity<-read.csv("activity.csv",header = T)
activity<-mutate(activity, date=as.POSIXct(date,format="%Y-%m-%d"))
```

## What is the mean total number of steps taken per day?

To analyse the number of steps per day, the activity data is grouped by day to summarise the number of steps taken. Note that days where no steps have been recorded have been removed from the dataset.

```{r}
byDay <- group_by(activity,date) %>%
    summarise(dailySteps = sum(steps, na.rm = T)) %>%
    filter(dailySteps != 0)
```

From this, we can derive the mean and median steps per day:

```{r eval=FALSE}
avgStepsPerDay <- mean(byDay$dailySteps)
medianStepsPerDay <- median(byDay$dailySteps)
```

    * The average number of steps per day is: `r mean(byDay$dailySteps)`  
    * The median number of steps per day is: `r median(byDay$dailySteps)`
  
A histogram of the recorded steps per day is shown below, to illustrate the relative frequency of steps per day. Note that this excludes days that had missing recorded values.

```{r}
#hist(byDay$dailySteps, xlab = "Number of steps per day", ylab="How many days", main = "Histogram of daily steps", breaks=20)
qplot(dailySteps, data=byDay, geom="histogram",  xlab = "Number of steps per day", ylab="How many days", main = "Histogram of daily steps")
```

## What is the average daily activity pattern?

To observe how the daily activity varies across the day, below is a plot showing the average steps taken for each 5 minute period averaged across all measured days.

```{r}
byInterval <- group_by(activity,interval) %>%
    summarise(avgSteps = mean(steps, na.rm = T)) %>%
    mutate(hm = ISOdatetime(1970,1,1,interval%/%60,interval%%60,0)) %>%
    mutate(hmc = strftime(hm,format="%H:%M"))
plot(byInterval$hm,byInterval$avgSteps,type="l", xlab="Time of day", ylab="Avg steps", main="Average activity by time of day")
```

```{r}
maxInterval <- byInterval[byInterval$avgSteps==max(byInterval$avgSteps),c("interval","hmc")]
```

As can be seen the interval that has the highest average number of steps is `r maxInterval[[1,"interval"]]` which equates to a time of day of `r maxInterval[[1,"hmc"]]`

## Inputing missing values



## Are there differences in activity patterns between weekdays and weekends?
